---
title: "Motif DB stats"
author: "Frank (zhangc518@gmail.com)"
date: "`r date()`"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
# 
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)

jaspar <- read.table("./db/jaspar2018_vertebrates.txt",skip = 1,stringsAsFactors = F)[,2:3] #579
tfclass <- readRDS("./db/tfclass.rds") 
dic<- readRDS("./db/dic_jaspar_tfclass.rds")
```

# jaspar stats 

The `JASPAR motifs`  were downloaded from [Jaspar](http://jaspar.genereg.net/downloads/) choosing `JASPAR2018_CORE_vertebrates_non-redundant` in `MEME` format. Then a customized function `getEnsemblFromJaspar` [code](./aux_funs.R) was used to get Ensembl ID from `Japsar id` via `uiportID` from Jaspar's record. 

```{r}
# number of TF motifs in Jaspar 
nrow(jaspar) 
# How many of them are protein complex? 
length(grep(":",jaspar$V3)) 
```

# tfclass stats 

[TFclass database](http://tfclass.bioinf.med.uni-goettingen.de/about.jsf) were downloaded (2018-05-14), [parsed](./scripts/parseTfclassTTL.sh) and [extacted](./scripts/tfclass.R) the human and mouse genes' Ensembl ID. 

```{r}
# How many unique genus (i.e. TF genes) in parsed TFClass?
length(unique(tfclass$merge$genus.id)) 

# How many subfamilies in parsed TFClass?
length(unique(tfclass$merge$subfamily.id))

# How many families in parsed TFClass?
length(unique(tfclass$merge$family.id))

```

# jaspar-tfclass dictionary stats 

`JASPAR` and `TFClass` were majorly [mapped](./cmpDB.R) by matching `Ensembl ID` except `20` TFs by `TF Name` directly. 

```{r}
# how many Jasper motif have annotation? 
a=nrow(dic$jasparTOtfclass %>% 
  dplyr::select(jaspar.id,genus.id) %>% 
  group_by(jaspar.id)%>%
  summarise(genus.id.2 = paste0(genus.id,collapse = ';')))
paste(a,nrow(jaspar),sep = "/")

# which jaspar motif don't have annotation?
noAnno <- subset(jaspar, V2 %in% setdiff(jaspar$V2,dic$jasparTOtfclass$jaspar.id))
noAnno.2 <- dic$jasparTOensembl[setdiff(jaspar$V2,dic$jasparTOtfclass$jaspar.id)]
knitr::kable(data.frame(jaspar.id=noAnno$V2,jaspar.name=noAnno$V3,
           ensembl.id= noAnno.2[noAnno$V2],
           url= paste0("http://jaspar2018.genereg.net/matrix/",noAnno$V2))%>%
  rowid_to_column('idx'))
```



