---
title: "Motif DB stats"
author: "Frank (zhangc518@gmail.com)"
date: "`r date()`"
always_allow_html: yes
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)
require(tidyverse)
require(knitr)
require(kableExtra)

jaspar <- read.table("./db/jaspar2018_vertebrates.txt",skip = 1,stringsAsFactors = F)[,2:3] #579
tfclass <- readRDS("./db/tfclass.rds") 
dic<- readRDS("./db/dic_jaspar_tfclass.rds")
```

# jaspar stats 

The `JASPAR motifs`  were downloaded from [Jaspar](http://jaspar.genereg.net/downloads/) choosing `JASPAR2018_CORE_vertebrates_non-redundant` in `MEME` format. Because there is no gene ids provided in the Jaspar database, the corresponding gene's `Ensembl ID` was fatched by querying the provided `UniprotID` from [UniprotID](https://www.uniprot.org/uniprot/) using a customized function `getEnsemblFromJaspar` [code](./aux_funs.R). 

The numbers are shown:
```{r}
kable(data.frame(
  n.jaspar=nrow(jaspar),
  n.complex=length(grep(":",jaspar$V3))))%>%
  kable_styling(c("striped", "bordered"),full_width = F)
```

# tfclass stats 

[TFclass database](http://tfclass.bioinf.med.uni-goettingen.de/about.jsf) were downloaded (2018-05-14), [parsed](./scripts/parseTfclassTTL.sh) and [extacted](./scripts/tfclass.R) the human and mouse genes' Ensembl ID. 

The number of each categories are shown: 

```{r}
kable(
  data.frame(n.genus=length(unique(tfclass$merge$genus.id)),
              n.subfamily=length(unique(tfclass$merge$subfamily.id)),
              n.family=length(unique(tfclass$merge$family.id))-1,
              n.class=length(unique(sub("\\.[0-9]+$","",unique(tfclass$merge$family.id)))),
              n.superclass=length(unique(sub("\\.[0-9]+\\.[0-9]+$","",unique(tfclass$merge$family.id))))-1))%>%
  kable_styling(c("striped", "bordered"),full_width = F) 
                         
```


There are two tfs have no family info in TFClass db:

```{r}
knitr::kable((tfclass$merge %>%
               filter(subfamily.id==""))[,3:4])%>%
  kable_styling(c("striped", "bordered"),full_width = F) 
```

# jaspar-tfclass dictionary stats 

`JASPAR` and `TFClass` were majorly [mapped](./cmpDB.R) by matching `Ensembl ID` except `20` TFs by `TF Name` directly. 

There are **`r paste(nrow(dic$jasparTOtfclass %>% dplyr::select(jaspar.id,genus.id) %>%   group_by(jaspar.id)%>%summarise(genus.id.2 = paste0(genus.id,collapse = ';'))),nrow(jaspar),sep = "/")`** jaspar motifs have annotation in TFclass.

The jaspar motif that doesn't have annotation are shown:

```{r}
noAnno <- subset(jaspar, V2 %in% setdiff(jaspar$V2,dic$jasparTOtfclass$jaspar.id))
noAnno.2 <- dic$jasparTOensembl[setdiff(jaspar$V2,dic$jasparTOtfclass$jaspar.id)]
knitr::kable(data.frame(jaspar.id=noAnno$V2,jaspar.name=noAnno$V3,
           ensembl.id= noAnno.2[noAnno$V2],
           url= paste0("http://jaspar2018.genereg.net/matrix/",noAnno$V2))%>%
  rowid_to_column('idx'))%>%
  kable_styling(c("striped", "bordered"),full_width = F) 
```

## Subfamily coverage

For TFclass, the number of jaspar motifs found for each subfamily ranged from 0 to 27. The detailed distribution is shown: 

```{r,echo=F,warning=F,message=F}
require(highcharter)
stats.jaspar<-dic$merged%>%
  group_by(subfamily.id)%>%
  summarise(n.jaspar=length(unique(jaspar.id[!is.na(jaspar.id)])))%>%
  complete(subfamily.id,fill = list(n.jaspar=0))%>%
  arrange(as.character(subfamily.id))
stats.jaspar<- stats.jaspar[-1,]
cnt <- count(stats.jaspar[-nrow(stats.jaspar),],n.jaspar)%>%
  rename(n.tf.subfamilty=n)

highchart() %>% 
  hc_chart(type = "column") %>% 
  hc_xAxis(categories=cnt$n.jaspar,
           title=list(text="n.jaspar"))%>%
  hc_yAxis(title=list(text="n.tf.subfamily"))%>%
  hc_add_series(data = cnt$n.tf.subfamilty)
```

There are 178 subfamilies have no jaspar motif:

```{r,echo=F,warning=F,message=F}
sf.noJaspar<-dic$merged%>% 
                filter(subfamily.id %in% stats.jaspar$subfamily.id[stats.jaspar$n.jaspar==0])%>%
                dplyr::select(-starts_with("jaspar"))%>%
                dplyr::select(-starts_with("ensem"))%>%
                group_by(subfamily.id)%>%
                summarise_all(funs(paste(unique(.),collapse = ";")))%>%
                arrange(subfamily.id)
DT::datatable(sf.noJaspar)
```

---
However, of these 178 subfamilies: 

```{r}
ids <- data.frame(undefined.sf=grepl("^0\\.",sf.noJaspar$subfamily.id),
           zinc_finger_more_than_3=grepl("^2\\.3\\.3\\.",sf.noJaspar$subfamily.id),
           dispersed_zinc_finger=grepl("^2\\.3\\.4\\.",sf.noJaspar$subfamily.id),
           skiped_sf=grepl("^[0-9]+\\.[0-9]+\\.[0-9]+\\.0$",sf.noJaspar$subfamily.id))%>%
  mutate(remain= !(undefined.sf|zinc_finger_more_than_3|dispersed_zinc_finger|skiped_sf))
ann<- c("0.x.x.x","2.3.3.x","2.3.4.x","x.x.x.0","rest")
cnt.2 <- data.frame(N=sapply(ids, sum),
                 id_type=ann)
kable(cnt.2)%>%
  kable_styling(c("striped", "bordered"),full_width = F) 
```

So in the end, there are **`r paste(cnt.2$N[5],sum(cnt$n.tf.subfamilty)-sum(cnt.2$N[-5]),sep="/")`** TFclass subfamily that doesn't contain a Jaspar motif. Below is the updated counts stats:

```{r}
cnt$n.tf.subfamilty[1]<-cnt.2$N[5]
highchart() %>% 
  hc_chart(type = "column") %>% 
  hc_xAxis(categories=cnt$n.jaspar,
           title=list(text="n.jaspar"))%>%
  hc_yAxis(title=list(text="n.tf.subfamily"))%>%
  hc_add_series(data = cnt$n.tf.subfamilty)

```


The jaspar and tfclass match can be searched here: 

```{r,echo=F,warning=F,message=F}
DT::datatable(
  stats.jaspar[stats.jaspar$n.jaspar>0,]%>% 
    right_join(dic$merged%>% 
                filter(subfamily.id %in% stats.jaspar$subfamily.id[stats.jaspar$n.jaspar>0])%>%
                dplyr::select(-starts_with("jaspar"))%>%
                dplyr::select(-starts_with("ensem"))%>%
                group_by(subfamily.id)%>%
                summarise_all(funs(paste(unique(.),collapse = ";"))))%>%
                arrange(subfamily.id)%>%
    drop_na())
```

## Family coverage


```{r}
stats.jaspar<-dic$merged%>%
  group_by(family.id)%>%
  summarise(n.jaspar=length(unique(jaspar.id[!is.na(jaspar.id)])))%>%
  complete(family.id,fill = list(n.jaspar=0))%>%
  arrange(as.character(family.id))
stats.jaspar<- stats.jaspar[-1,]
stats.jaspar<-stats.jaspar%>% filter(!(grepl("^0\\.",family.id)|
                            family.id=="2.3.3" |
                           family.id=="2.3.4"))
cnt <- count(stats.jaspar[-nrow(stats.jaspar),],n.jaspar)%>%
  rename(n.tf.family=n)

highchart() %>% 
  hc_chart(type = "column") %>% 
  hc_xAxis(categories=cnt$n.jaspar,
           title=list(text="n.jaspar"))%>%
  hc_yAxis(title=list(text="n.tf.family"))%>%
  hc_add_series(data = cnt$n.tf.family)
```

There are **`r sum(cnt$n.tf.family)-cnt$n.tf.family[1]`** out of **`r sum(cnt$n.tf.family)`** have at least one Jaspar motif found. Note: the total number of family here are counted after filtered out those `undefined`, `zinc_finger_more_than_3`, and `dispersed_zinc_finger`. 